---
title: "Lab 3"
#author: "Jefferson Ratliff"
date: "Math 241, Week 3"
output:
  pdf_document
urlcolor: blue
---

```{r setup, include=FALSE}
# Do not modify this chunk.
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
libs <- c('tidyverse','knitr','viridis', 'mosaic','mosaicData','babynames', 'Lahman','nycflights13','rnoaa')
for(l in libs){
  if(!require(l,character.only = TRUE, quietly = TRUE)){
    message( sprintf('Did not have the required package << %s >> installed. Downloading now ... ',l))
    install.packages(l) 
  }
  library(l, character.only = TRUE, quietly = TRUE)
}
```



## Due: Friday, February 16th at 8:30am

## Goals of this lab

1. Practice creating functions.
1. Practice refactoring your code to make it better!  Therefore for each problem, make sure to test your functions.


### Problem 1: Subset that R Object

Here are the R objects we will use in this problem (`dats`, `pdxTreesSmall` and `ht`).    

```{r}
library(pdxTrees)
library(mosaicData)
library(ggplot2)

pdxTrees <- get_pdxTrees_parks()
# Creating the objects
dats <- list(pdxTrees  = head(pdxTrees),
             Births2015 = head(Births2015),
             HELPrct = head(HELPrct), 
             sets = c("pdxTrees", "Births2015", 
                      "HELPrct"))

pdxTreesSmall  <- head(pdxTrees)

ht <- head(pdxTrees$Tree_Height, n = 15)
```

a. What are the classes of `dats`, `pdxTreesSmall` and `ht`?

```{r}
class(dats)
class(pdxTreesSmall)
class(ht)
```

<<<<<<< Updated upstream

=======
>>>>>>> Stashed changes
b. Find the 10th, 11th, and 12th values of `ht`.

```{r}
ht[10:12]
```


c. Provide the `Species` column of `pdxTrees` as a data frame with one column.

```{r}
pdxTrees_species_df <- select(pdxTrees, Species)
<<<<<<< Updated upstream

pdxTrees_species_df
=======
>>>>>>> Stashed changes
```


d. Provide the `Species` column of `pdxTrees` as a character vector.

```{r}
pdxTrees_species_chr <- head(pdxTrees$Species)
as.character(pdxTrees_species_chr)
```


e. Provide code that gives us the second entry in `sets` from `dats`.

```{r}
dats$sets[2]
```


f. Subset `pdxTreesSmall` to only `Douglas-fir` and then provide the `DBH` and `Condition` of the 4th `Douglas-fir` in the dataset.  (Feel free to mix in some `tidyverse` code if you would like to.)

```{r}
<<<<<<< Updated upstream
library(dplyr)

pdxTreesSmall %>%
  filter(Species == "Douglas-fir") %>%
  slice(4) %>%
  select(DBH, Condition)
```


=======
pdxTreesSmall <- filter(pdxTreesSmall, Common_Name == "Douglas-Fir")
pdxTreesSmall$DBH[4]
pdxTreesSmall$Condition[4]
```
>>>>>>> Stashed changes

### Problem 2: Function Creation

Figure out what the following code does and then turn it into a function. For your new function, do the following:

* Test it.
* Provide default values (when appropriate).
* Use clear names for the function and arguments.
* Make sure to appropriately handle missingness.
* Generalize it by allowing the user to specify a confidence level.
* Check the inputs and stop the function if the user provides inappropriate values.


```{r}
library(pdxTrees)
thing1 <- length(pdxTrees$DBH)
thing2 <- mean(pdxTrees$DBH)
thing3 <- sd(pdxTrees$DBH)/sqrt(thing1)
thing4 <- qt(p = .975, df = thing1 - 1)
thing5 <- thing2 - thing4*thing3
thing6 <- thing2 + thing4*thing3
```


```{r}
calculate_CI <- function(data, confidence_level = 0.95) {
  
<<<<<<< Updated upstream
=======
  # Check if data is provided
>>>>>>> Stashed changes
  if (missing(data) || is.null(data)) {
    stop("Please provide a valid data set.")
  }
  
<<<<<<< Updated upstream

  if (confidence_level <= 0 || confidence_level >= 1) {
    stop("Confidence level must be between 0 and 1.")
  }


  n <- length(data)
  mean_val <- mean(data)
  std_err <- sd(data) / sqrt(n)
  t_value <- qt(p = (1 + confidence_level) / 2, df = n - 1)

  lower_bound <- mean_val - t_value * std_err
  upper_bound <- mean_val + t_value * std_err
  
  
  result <- list(
    mean = mean_val,
    lower_bound = lower_bound,
    upper_bound = upper_bound,
    confidence_level = confidence_level
  )
  
  return(result)
}

result <- calculate_CI(pdxTrees$DBH, confidence_level = 0.95)
print(result)
```
=======
  # Check if confidence level is between 0 and 1
  if (confidence_level <= 0 || confidence_level >= 1) {
    stop("Confidence level must be between 0 and 1.")
  }
>>>>>>> Stashed changes

  # Calculate necessary values
  n <- length(data)
  mean_val <- mean(data)
  std_err <- sd(data) / sqrt(n)
  t_value <- qt(p = (1 + confidence_level) / 2, df = n - 1)
  
  # Calculate confidence interval
  lower_bound <- mean_val - t_value * std_err
  upper_bound <- mean_val + t_value * std_err
  
  # Return a named list with results
  result <- list(
    mean = mean_val,
    lower_bound = lower_bound,
    upper_bound = upper_bound,
    confidence_level = confidence_level
  )
  
  return(result)
}

# Test the function with pdxTrees$DBH

result <- calculate_CI(pdxTrees$DBH, confidence_level = 0.95)
print(result)
```


### Problem 3: Wrapper Function for your `ggplot`


While we (i.e. Math 241 students) all love the grammar of graphics, not everyone else does. So for this problem, we are going to practice creating wrapper functions for `ggplot2`.

Here's an example of a wrapper for a histogram.  Notice that I can't just list the variable name as an argument.  The issue has to do with how many of the `tidyverse` functions evaluate the arguments. Therefore we have to quote (`enquo()`) and then unquote (`!!`) the arguments.  (If you want to learn more, go [here](https://tidyeval.tidyverse.org/).)

```{r}
ggplot(data = pdxTrees, mapping = aes(x = DBH)) +
  geom_histogram()

histo <- function(data, x){
   x <- enquo(x)
  ggplot(data = data, mapping = aes(x = !!x)) +
    geom_histogram()
}

histo(pdxTrees, DBH)
```

a. Edit `histo()` so that the user can set 

* The number of bins
* The fill color for the bars
* The color outlining the bars

```{r}
histo <- function(data, x, bins = 30, fill_color = "blue", outline_color = "black") {
  x <- enquo(x)
  ggplot(data = data, mapping = aes(x = !!x)) +
    geom_histogram(bins = bins, fill = fill_color, color = outline_color)
}

<<<<<<< Updated upstream
# Test it with custom parameters
histo(pdxTrees, DBH, bins = 20, fill_color = "green", outline_color = "red")
```


b. Write code to create a basic scatterplot with `ggplot2`. Then write and test a function to create a basic scatterplot.

```{r}
library(ggplot2)


basic_scatterplot <- ggplot(data = mtcars, aes(x = mpg, y = wt)) +
  geom_point()


print(basic_scatterplot)


=======
histo(pdxTrees, DBH, bins = 20, fill_color = "green", outline_color = "red")
```

b. Write code to create a basic scatterplot with `ggplot2`. Then write and test a function to create a basic scatterplot.

```{r}
ggplot(data = mtcars, aes(x = mpg, y = wt)) +
  geom_point()

>>>>>>> Stashed changes
scatterplot_basic <- function(data, x, y) {
  ggplot(data = data, aes(x = {{ x }}, y = {{ y }})) +
    geom_point()
}

<<<<<<< Updated upstream

=======
>>>>>>> Stashed changes
scatterplot_basic(mtcars, mpg, wt)
```

c. Modify your scatterplot function to allow the user to ...

* Color the points by another variable.
* Set the transparency.

```{r}
<<<<<<< Updated upstream
scatter_function <- function(data, x, y, alpha = 0.5, color = Condition){
   x <- enquo(x)
   y <- enquo(y)
   color <- enquo(color)
  ggplot(data = data, mapping = aes(x = !!x, y = !!y, color = !!color)) +
    geom_point(alpha = alpha)
}

# Testing scatter function
scatter_function(pdxTrees, Tree_Height, Carbon_Storage_value, alpha = 0.9, color = Native)
=======
scatter <- function(data, x, y, color_var = NULL, alpha = 1) {
  x <- enquo(x)
  y <- enquo(y)
  color_var <- enquo(color_var)
  
  ggplot(data = data, aes(x = !!x, y = !!y, color = !!color_var)) +
    geom_point(alpha = alpha)
}

scatter(pdxTrees, x = DBH, y = Tree_Height, color_var = Species, alpha = 0.7)
>>>>>>> Stashed changes
```


d. Write and test a function for your favorite `ggplot2` graph.

```{r}
<<<<<<< Updated upstream
library(ggplot2)

#Basic violin plot code
ggplot(data = pdxTrees, aes(x = Condition, y = Carbon_Storage_lb)) +
  geom_violin()
```

```{r}
#Function for basic violin plot
violin_function <- function(data, x, y){
   x <- enquo(x)
   y <- enquo(y)
  ggplot(data = data, mapping = aes(x = !!x, y = !!y)) +
    geom_violin()
}

#Testing violin function
violin_function(pdxTrees, Condition, Carbon_Storage_lb)
=======
bar_plot <- function(data, x, y, fill_color = "skyblue", outline_color = "black") {
  x <- enquo(x)
  y <- enquo(y)
  
  ggplot(data = data, mapping = aes(x = !!x, y = !!y)) +
    geom_bar(stat = "identity", fill = fill_color, color = outline_color)
}

bar_plot(pdxTrees, Species, n_distinct(Tree_Height), fill_color = "orange", outline_color = "brown")
>>>>>>> Stashed changes
```

### Problem 4: Functioning `dplyr`


a. Take the following code and turn it into an R function to create a **conditional proportions** table.  Similar to `ggplot2`, you will need to quote and unquote the variable names.  Make sure to test your function!

```{r}
conditional_proportions_table <- function(data, group_var, count_var) {
  group_var <- enquo(group_var)
  count_var <- enquo(count_var)
  
  data %>%
    count(!!group_var, !!count_var) %>%
    group_by(!!group_var) %>%
    mutate(prop = n / sum(n)) %>%
    ungroup()
}

conditional_proportions_table(pdxTrees, Native, Condition)
```
```{r}
conditional_proportions_table <- function(data, var1, var2) {
  data %>%
    count(!!enquo(var1), !!enquo(var2)) %>%
    group_by(!!enquo(var1)) %>%
    mutate(prop = n/sum(n)) %>%
    ungroup()
}

result <- conditional_proportions_table(pdxTrees, Native, Condition)
print(result)
```


b.  Write a function to compute the mean, median, sd, min, max, sample size, and number of missing values of a quantitative variable by the categories of another variable.  Make sure the output is a data frame (or tibble).  Don't forget to test your function.

```{r}
<<<<<<< Updated upstream
library(dplyr)

quant_summary_by_category <- function(data, category_var, quant_var) {
  data %>%
    group_by(!!enquo(category_var)) %>%
    summarize(
      mean = mean(!!enquo(quant_var), na.rm = TRUE),
      median = median(!!enquo(quant_var), na.rm = TRUE),
      sd = sd(!!enquo(quant_var), na.rm = TRUE),
      min = min(!!enquo(quant_var), na.rm = TRUE),
      max = max(!!enquo(quant_var), na.rm = TRUE),
      n = n(),
      missing_values = sum(is.na(!!enquo(quant_var)))
=======
summary_by_category <- function(data, category_var, quant_var) {
  category_var <- enquo(category_var)
  quant_var <- enquo(quant_var)
  
  data %>%
    group_by(!!category_var) %>%
    summarise(
      mean = mean(!!quant_var, na.rm = TRUE),
      median = median(!!quant_var, na.rm = TRUE),
      sd = sd(!!quant_var, na.rm = TRUE),
      min = min(!!quant_var, na.rm = TRUE),
      max = max(!!quant_var, na.rm = TRUE),
      n = n(),
      na_count = sum(is.na(!!quant_var))
>>>>>>> Stashed changes
    ) %>%
    ungroup()
}

<<<<<<< Updated upstream

example_data <- data.frame(
  Category = rep(c("A", "B", "C"), each = 5),
  Value = c(1, 2, 3, 4, NA, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14)
)

result <- quant_summary_by_category(example_data, Category, Value)
print(result)

```

=======
```

```{r}
summary_result <- summary_by_category(pdxTrees, Native, DBH)
print(summary_result)
```


>>>>>>> Stashed changes

### Problem 5: another `babynames` exercise

Write a function called grab_name that, when given a **name** ***and a year*** as an argument, returns the rows from the `babynames` data frame in the `babynames` package that match that name for that year (and returns an error if that name and year combination does not match any rows). Run the function once with the arguments **Ezekiel and 1883** and once with **Ezekiel and 1983**.

```{r, eval=TRUE}
grab_name <- function(myname, myyear) {
<<<<<<< Updated upstream
  result <- babynames %>% 
    filter(name == myname, year == myyear)
  
  if (nrow(result) == 0) {
    stop("No matching rows found for the given name and year combination.")
=======
  library(babynames)
  
  result <- babynames %>%
    filter(name == myname, year == myyear)
  
  if (nrow(result) == 0) {
    stop("No matching rows found for the given name and year.")
>>>>>>> Stashed changes
  }
  
  return(result)
}

<<<<<<< Updated upstream
# Test with Ezekiel and 1883
result_1883 <- grab_name("Ezekiel", 1883)
print(result_1883)

# Test with Ezekiel and 1983
result_1983 <- grab_name("Ezekiel", 1983)
print(result_1983)
=======

grab_name("Ezekiel", 1883)


grab_name("Ezekiel", 1983)
>>>>>>> Stashed changes
```